---
title: "Brushing model 1"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

# Objective

We want to analyze variations in the amount of time spent brushing each dental region,
session-to-session and among participants.

In particular, we want to know whether the participants spent more time brushing their lower teeth
than upper teeth, left versus right, and inner (lingual) versus outer (buccal) surfaces.

```{r, "Load data", include = FALSE}

library(magrittr)
library(dplyr)
library(tidyr)
library(papeR)
library(brushing.habits)
library(lme4)
library(plotly)
library(pander)
library(broom.mixed)
library(glmmTMB)
library("bbmle")
library(sjPlot)
library(parameters)

panderOptions('digits', 3)
panderOptions('table.split.table', Inf)

```

# Model for brushing duration per region

```{r}

fit0 = glmmTMB(
  family = "poisson",
  data  = region_durations,
  formula = n_samples ~ Surface + Side + Jaw + (1 + Surface + Jaw + Side | Participant) + (1 | Participant:Session),
  ziformula = ~ 0
)

fit1 = update(fit0, ziformula = ~ 1)
fit2 = update(fit1, ziformula = ~ Surface + Side + Jaw)
fit3 = update(fit2, ziformula = ~ Surface + Side + Jaw + (1 | Participant))
fit4 = update(fit3, ziformula = ~ Surface + Side + Jaw + (1 | Participant / Session))

# doesn't converge:
# fit5 = update(fit3, ziformula = ~ Surface + Side + Jaw + (1 + Surface + Jaw + Side | Participant) + (1 | Participant / Session))


fit3nb = update(fit3, family = nbinom2)
fit4nb = update(fit4, family = nbinom2)
fit5nb = update(fit4nb, dispformula = ~ 0 + factor(Participant))


aics = AICtab(
  fit3nb, fit4nb, fit5nb,
  fit0, fit1, fit2,
  fit3,
  fit4)

bics = BICtab(
  fit3nb, fit4nb,fit5nb,
  fit0, fit1, fit2,
  fit3,
  fit4)

region_chosen = fit5nb

aic_chosen = AIC(region_chosen)
aic_poisson = AIC(fit4)
bic_chosen =  BIC(region_chosen)
bic_poisson = BIC(fit4)


diff_aic = aic_poisson - aic_chosen
diff_bic = bic_poisson - bic_chosen

# The AIC for this model was `r format(aic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_aic, scientific = Inf, big.mark = ",")` less than a zero-inflated Poisson model with the same fixed and random effects. The BIC for this model was `r format(bic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_bic, scientific = Inf, big.mark = ",")` less than the zero-inflated Poisson model with the same fixed and random effects.
```

Here we fit a zero-inflated negative binomial regression model for the amount of time (measured in counts of 25 Hz samples) spent brushing each region, 
with a log-link for the count submodel and a logistic link for the zero-inflation submodel.
Both submodels had fixed effects for tooth surface, mouth side, and jaw,
and random effects on the intercept by session nested in participant, to account for person-to-person and session-to-session variability in overall brushing duration.
The count submodel also included participant-specific random effects for tooth surface, mouth side, and jaw.
We attempted to add participant-specific random effects for tooth surface, mouth side, and jaw in the zero-inflation submodel, but the estimation algorithm failed to converge for that extended model. 

The AIC for this model was `r format(aic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_aic, scientific = Inf, big.mark = ",")` less than a zero-inflated Poisson model with the same fixed and random effects. The BIC for this model was `r format(bic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_bic, scientific = Inf, big.mark = ",")` less than the zero-inflated Poisson model with the same fixed and random effects.

## Parameter estimates:


```{r,results = "asis"}

# a = tab_model(fit4, show.icc = FALSE, show.re.var = TRUE, show.se = TRUE, show.ngroups = FALSE, show.obs = FALSE, show.r2 = FALSE, transform = NULL)
# coef(summary(fit4))$cond %>%
#   format_fixef_table %>%
#   pander()

# this argument is used in parameters:::.prepare_x_for_print
to_print = c("Parameter", "Coefficient", 
             "SE",
             "CI", "CI_low", "CI_high", "p")
parameters(region_chosen) %>% print_md(select = to_print)

# VarCorr(fit0)
```

### Estimates of participant-level random effects for count submodel:

```{R}

region_chosen %>% format_ranef_table(comp = "cond", combine_stats = FALSE) %>% pander()

```

### Estimates of participant-level random effects for zero-inflation submodel:

```{R}

region_chosen %>% format_ranef_table(comp = "zi", combine_stats = FALSE) %>% pander()

```


# Brushing duration with excessive pressure

Again, we fit a zero-inflated negative binomial model; both the zero-inflation and count submodels include 
fixed effects for tooth surface, side, and jaw, and random effects on the intercept by participant and session.
An extended model with participant-level random effects on tooth surface, side, and jaw failed to converge.

```{r}

# this creates a convergence problem:
# pressure_durations %<>% pressure_durations  %>% mutate(Surface = relevel(factor(Surface), ref = "Occlusal"))
pressure1 = glmmTMB(
  family = nbinom2,
  data  = pressure_durations,
  # formula = n_samples ~ Surface + Side + Jaw + (1 + Surface + Jaw + Side | Participant) + (1 | Session:Participant),
  formula = n_samples ~ Surface + Side + Jaw + (1 | Participant / Session),
  ziformula = ~ Surface + Side + Jaw + (1 | Participant / Session),
  dispformula = ~ 1)

pressure_pois = glmmTMB(
  family = poisson,
  data  = pressure_durations,
  # formula = n_samples ~ Surface + Side + Jaw + (1 + Surface + Jaw + Side | Participant) + (1 | Session:Participant),
  formula = n_samples ~ Surface + Side + Jaw + (1 | Participant / Session),
  ziformula = ~ Surface + Side + Jaw + (1 | Participant / Session))


aic_chosen = AIC(pressure1)
aic_poisson = AIC(pressure_pois)
bic_chosen =  BIC(pressure1)
bic_poisson = BIC(pressure_pois)


diff_aic = aic_poisson - aic_chosen
diff_bic = bic_poisson - bic_chosen

# The AIC for this model was `r format(aic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_aic, scientific = Inf, big.mark = ",")` less than a zero-inflated Poisson model with the same fixed and random effects. The BIC for this model was `r format(bic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_bic, scientific = Inf, big.mark = ",")` less than the zero-inflated Poisson model with the same fixed and random effects.
```

The AIC for this model was `r format(aic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_aic, scientific = Inf, big.mark = ",")` less than a zero-inflated Poisson model with the same fixed and random effects. The BIC for this model was `r format(bic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_bic, scientific = Inf, big.mark = ",")` less than the zero-inflated Poisson model with the same fixed and random effects.

## Parameter Estimates

```{r}

parameters(pressure1) %>% print_md(select = to_print)
```


### Estimates of participant-level random effects for count submodel:

```{R}

pressure1 %>% format_ranef_table(comp = "cond", combine_stats = FALSE) %>% pander()

```

### Estimates of participant-level random effects for zero-inflation submodel:

```{R}

pressure1 %>% format_ranef_table(comp = "zi", combine_stats = FALSE) %>% pander()

```

# Total duration of brushing

Here we fit a negative binomial regression model for the total duration of each brushing session (again measured in counts of 25 Hz samples),
with a log-link, random intercepts by participant, and person-specific dispersion parameters (modeled as fixed effects due to the limitations of the modeling package).

```{r}

td_poisson = glmmTMB(
  family = poisson,
  data  = session_durations2,
  formula = n_samples ~ (1 | Participant))

td_poisson2 = glmmTMB(
  family = poisson,
  data  = session_durations2,
  formula = n_samples ~ (1 | Participant / Session))


td_nbinom1 = glmmTMB(
  family = nbinom2,
  data  = session_durations2,
  formula = n_samples ~ (1 | Participant),
  dispformula = ~ 1)

td_nbinom2 = glmmTMB(
  family = nbinom2,
  data  = session_durations2,
  formula = n_samples ~ (1 | Participant),
  dispformula = ~ factor(Participant) + 0)

aics = AICtab(td_poisson, td_poisson2, td_nbinom1, td_nbinom2)
bics = BICtab(td_poisson, td_poisson2, td_nbinom1, td_nbinom2)

chosen_model = td_nbinom2


aic_chosen = AIC(chosen_model)
aic_poisson = AIC(td_poisson)
bic_chosen =  BIC(chosen_model)
bic_poisson = BIC(td_poisson)


diff_aic = aic_poisson - aic_chosen
diff_bic = bic_poisson - bic_chosen

# The AIC for this model was `r format(aic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_aic, scientific = Inf, big.mark = ",")` less than a zero-inflated Poisson model with the same fixed and random effects. The BIC for this model was `r format(bic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_bic, scientific = Inf, big.mark = ",")` less than the zero-inflated Poisson model with the same fixed and random effects.

```

The AIC for this model was `r format(aic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_aic, scientific = Inf, big.mark = ",")` less than a zero-inflated Poisson model with the same fixed and random effects. The BIC for this model was `r format(bic_chosen, scientific = Inf, big.mark = ",")`, which was `r format(diff_bic, scientific = Inf, big.mark = ",")` less than the zero-inflated Poisson model with the same fixed and random effects.

## Parameter estimates

```{r}
parameters(chosen_model) %>% 
  print_md(select = to_print)

```


### Estimates of participant-level random effects:

```{R}

chosen_model %>% format_ranef_table(comp = NULL, combine_stats = FALSE) %>% pander()

```

Here is a table of participant-specific estimates:

```{r}
### within-person variability

temp = 
  session_durations2 %>% 
  group_by(Participant) %>% 
  summarise(
    `sample mean (seconds)` = mean(n_samples) / 25,
    `sample var (seconds^2)` = var(n_samples) / 25^2,
    `sample sd (seconds)` = sd(n_samples) / 25) %>%
  mutate(
    `lambda (samples)` = exp(coef(chosen_model)$cond$Participant$`(Intercept)`),
    `lambda (seconds)` = `lambda (samples)`/ 25,
    kappa = exp(fixef(chosen_model)$disp),
    `model var (seconds^2)` = (`lambda (samples)` * (1 + `lambda (samples)` / kappa)) / 25^2,
    `model sd (seconds)` = sqrt(`model var (seconds^2)`)
  )

pander(temp)
```

